---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

This is a simple [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code. 

## Chicago Weather from weather stations

The  URLs for data are on the portal https://data.cityofchicago.org/ 

In particular the hourly data from the three beach weather stations has a Base URL: 
https://data.cityofchicago.org/resource/7edu-s3u7.json 
The API documentation for this dataset is at:
https://dev.socrata.com/foundry/data.cityofchicago.org/77jv-5zb8

Let's read the dataset and display the 'head' of it:

```{r}
df <- read.csv("https://data.cityofchicago.org/resource/7edu-s3u7.csv")
head(df)

```

Unfortunately the web-service only lets you download 1000 observations in this way and displays the Date/Time with AM/PM which is near impossible to deal with in R.

As we can see there are three weather stations posting observations every hour as a row (a separate row for each of them). It is just logical to filter out the observations for each of the stations from the dataset with a 'where' query (but we will need to load the library for that) In particular:

```{r}
library("RSocrata", lib.loc="~/R/win-library/3.3")

fdf <- read.socrata("https://data.cityofchicago.org/resource/7edu-s3u7.csv?$where=station_name=\"Foster Weather Station\"")
```

THis query read 6469 observations of 18 variables. Let's do the same with the other two stations.

```{r}
odf <- read.socrata("https://data.cityofchicago.org/resource/7edu-s3u7.csv?$where=station_name=\"Oak Street Weather Station\"")

sdf <- read.socrata("https://data.cityofchicago.org/resource/7edu-s3u7.csv?$where=station_name=\"63rd Street Weather Station\"")
```

Let's backup this data as *.rda files... just in case.

```{r}
save(fdf, file = "fdf.rda")
save(odf, file = "odf.rda")
save(sdf, file = "sdf.rda")
```

Also, let's download everything that can be downloaded directly from the web as a *.csv and re-save it in Excel, in case some weird formatting will get in the way. We will also have a combined data frame then and will be able to assign a category to a weather station and use it later in graphics.
It is in the directory, stored as: "Beach_Weather_Stations_-_Automated_Sensors_-_2016_-_Humidity.csv"
Downloaded and added the same for 2015 and the "Beach_Water_and_Weather_Sensor_Locations_-_Map.csv" file with the locations of the sensors.

Now we can attempt plotting a little bit

```{r}
library("ggplot2", lib.loc="~/R/win-library/3.3")

qplot(Measurement.Timestamp, Air.Temperature, data = fdf, color = I("magenta"), 
      size = I(.5), alpha = I(1/2))
```


As we can see ggplot2 'understands' our Date/Time format. Interestingly enough there is a gap in data in May.

```{r}
qplot(Measurement.Timestamp, Air.Temperature, data = sdf, color = I("red"), 
      size = I(.5), alpha = I(1/2))
```

```{r}
qplot(Measurement.Timestamp, Air.Temperature, data = odf, color = I("green"), 
      size = I(.5), alpha = I(1/2))
```

The direct observation of data shows that indeed the gap (starting after May 9, 13:00 and until May 19 10am) exists.

All together Now!

```{r}
#df will be re-used for complete data set
df <- read.csv("Beach_Weather_Stations_-_Automated_Sensors_-_2016_-_Humidity.csv")
qplot(Measurement.Timestamp, Air.Temperature, data = df, colour = Station.Name)
```

Not exactly OK with the Dates read from the *.csv file.


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
